### 2015-04-07 conclusion.tex
Initial outline

%!TEX root = paper.tex

\section{Conclusion}
\label{sec:conclusion}

\todo{.5 pg}