### 2015-04-07 usecase-combinatoric.tex

%!TEX root = paper.tex

\subsection{Combinatoric Names}
\label{sec:usecase-combinatoric}

\subsubsection{Gperftools packaging}
\todo{.5 page}

\begin{verbatim}
	- combinatoric nightmare
	- compilers x versions
\end{verbatim}

\subsubsection{{\tt mpileaks} packaging}
\todo{.5 page}
\begin{verbatim}
	- compiler x version x MPI x mpi version
\end{verbatim}