### 2015-04-07 usecases.tex

%!TEX root = paper.tex

\section{Use Cases}
\label{sec:usecases}

\input{usecase-combinatoric}
\input{usecase-python}
\input{usecase-policy}