### 2015-04-09 abstract.tex
Updated build.

%!TEX root = paper.tex

The largest HPC centers spend considerable time supporting software configurations for thousands of users, but the complexity of HPC software is quickly outpacing the capabilities of existing software management tools. Typical scientific applications require specific versions of many dependency libraries, and they must build with many compilers and MPI versions. This configuration space is exponential in size, and it is infeasible to maintain a single software stack for all users at a site.  We present Spack, a novel package management tool that allows many software configurations to be built on demand. Spack provides a novel, recursive specification syntax, and it allows arbitrarily many software and dependency combinations to coexist on the same system. Spack's build mechanism ensures that each package can find its dependencies, regardless of a user's runtime environment settings. We show through several real-world use cases that Spack brings order to the chaos of HPC software installation management.