### 2015-04-16 implementation-virtual-deps.tex
updates through specs.

%!TEX root = paper.tex

\subsection{Versioned Virtual Dependencies}
	\todo{.5 page}