### 2015-04-16 implementation.tex

%!TEX root = paper.tex

\section{The Spack Package Manager}
\label{sec:implementation}
Based on our experiences at LLNL, we have developed
{\it Spack}, the Supercomputing PACKage manager.
Spack is written in Python.  We chose Python because it is
very widely used int he HPC community, and because it provides
flexible scripting capabilities. 
%
Like some prior systems, Spack supports arbitrarily many software installations.
It borrows ideas from Nix; packages are installed in their own prefixes,
and configurations are managed using hashes.

\noindent
Spack adds several essential capabilities for HPC:
\begin{enumerate}
\item {\bf Parametric builds} address the combinatorial problem:
      Packages are parameterized by name, version, architecture, compiler, 
      options, and dependencies.
\item To manage the configuration space, spack provides a novel, 
      recursive {\bf spec syntax} for dependency graphs.
\item Specs support versioned, ABI-incompatible interfaces like MPI through
      {\bf versioned virtual dependencies}.
\item Spack builds with {\bf compiler wrappers} that reduce the burden of build
      consistency on the package author.
\end{enumerate}

\input{implementation-packages}
\input{implementation-specs}
\input{implementation-virtual-deps}
\input{implementation-concretization}

\input{implementation-environment}