### 2015-04-17 implementation-virtual-deps.tex

%!TEX root = paper.tex

\subsection{Versioned Virtual Dependencies}

\begin{figure}
%
\begin{subfigure}{\linewidth}
	\begin{minted}[numbersep=5pt,
	              fontsize=\scriptsize,
	              frame=lines,
                  framesep=2mm]{python}
class Mvapich2(Package):    
    provides('mpi@:2.2', when='@1.9')
    provides('mpi@:3.0', when='@2.0')
\end{minted}
	\caption{%
		{\tt mvapich2} provides {\tt mpi} interface with specific versions.
		\label{fig:virtual-deps-provides}
	}
\end{subfigure}
%
\begin{subfigure}{\linewidth}
	\begin{minted}[numbersep=5pt,
	              fontsize=\scriptsize,
	              frame=lines,
                  framesep=2mm]{python}
class Mpileaks(Package):    
    depends_on('mpi')
\end{minted}
	\caption{%
		{\tt mpileaks} depends on any MPI version.
		\label{fig:virtual-deps-mpileaks}
	}
\end{subfigure}
%
\begin{subfigure}{\linewidth}
	\begin{minted}[numbersep=5pt,
	              fontsize=\scriptsize,
	              frame=lines,
                  framesep=2mm]{python}
class Gerris(Package):
    depends_on('mpi@2:')
\end{minted}
	\caption{%
		Gerris CFD solver requires MPI at version 2 or higher.
		\label{fig:virtual-deps-depends}
	}
\end{subfigure}
%
\caption{
	Virtual dependencies.
	\label{fig:virtual-deps}
}
\end{figure}

Many libraries share a common source interface and can be interchanged for
one another in a build.
The archetypal example in HPC is the MPI interface, which has a number of
open source implementations (MPICH, OpenMPI, MVAPICH, etc.) and vendor-specific
implementations. An application that can be built with one MPI can generally be
built with another.  BLAS implementations such as ATLAS, LAPACK-BLAS, MKL, etc.
are another common example.

At LLNL, we frequently need to build tools with many versions of MPI to support
the many different applications that are run at our center.
However, MPI has no ABI, so applications cannot simply be re-linked;
they must be recompiled and reinstalled. 
Complicating matters, the MPI interface is versioned, and some
packages need later versions of MPI to run correctly.  Often, MPI
{\it implementation} versions do not correspond obviously to MPI
interface versions, and determining the right version of 
MPI to pair with an application can be tedious and tricky.

To allow rapid composition of libraries by interface, Spack supports
{\it virtual dependencies}.  A virtual dependency is an abstract name
representing a library interface or capability instead of a library 
implementation.  Packages that need this capability need not depend on 
a specific implementation; they can depend on the virtual name, for which
the user or Spack can select an an implementation at build time.
Other package managers support the notion of virtual dependencies, but Spack
adds {\it versioning} to its interfaces.  This allows concepts like MPI
interface versions and BLAS levels to be represented directly, and Spack can
handle the details of managing complex constraints while the user focuses on 
composing software.

Figure~\ref{fig:virtual-deps-provides} shows how packages can provide
virtual interfaces in Spack.  The spec syntax is used to concisely
associate a range of {\tt mpi} versions (up to 2.2) with Mvapich version 1.9.
In Figure~\ref{fig:virtual-deps-mpileaks}, we see that {\tt mpileaks} requires
MPI but is unconstrained on version.  Since this spec satisfies the constraints in 
the {\tt mvapich2} package, {\tt Mvapich2} can be used to satisfy the {\tt mpi}
dependency in {\tt mpileaks}.  Similarly, the Gerris CFD library needs MPI
at version 2 or higher -- either {\tt mvapich2} 1.9 or 2.0 can be used to satisfy
this constrained dependency.