### 2015-04-18 usecases.tex
flow.

%!TEX root = paper.tex

\section{Use Cases}
\label{sec:usecases}

We have recently begun deploying the Spack system in production at
Livermore Computing

\input{usecase-combinatoric}
\input{usecase-python}
\input{usecase-policy}