### 2015-04-18 usecase-python.tex

%!TEX root = paper.tex

\subsection{Support for extensible languages}
\label{sec:usecase-python}

Python is becoming increasingly popular for HPC applications, 
both due to its flexibility as a language and due to its excellent support
for calling into native numerical code and achieving high performance. 
Python applications frequently make native calls into functions in
compiled numerical libraries like NumPy, FFTW, ATLAS, and other linear
algebra libraries.

LC maintains a number of Python installations for production application teams,
and maintenance of these repositories has been increasingly complex for
some time.  LC must maintain different Python installations containing different
versions of Python modules for different code teams, and the modules must be able
to make calls into optimized native libraries, for maximum performance.
Python itself was easy to package with Spack; we built a package for it like any
other and it compiled and ran

they must also call into high performance math libraries. Python installations
suffer from many of the environment problems we mentioned in Sections~\ref{env-rpath}.

Python

 \cite{eby:setuptools}

Python's package management

To support code teams using Python, LC maintains a number of

both for rapid prototyping

- LC maintains python installations for existing codes.

- maintain separate instances for different code teams
	- each code team needs to have particular python packages installed
	- need fast builds of math libraries to go with python installs

- model for python package installation doesn't really fit spack

- implemented best of both worlds solution:
	- combine spack's flexibility with python meta-package-manager convenience

	- many other languages fit this model; could use it w/them.

\todo{1 pg}

\begin{verbatim}
- need for special python handling (prefix issues)
- solution of offering "meta package management"
- activate/deactivate python packages
- installs are in own prefix but can be merged into extended prefix
- design is extensible, allows for other anguages to use custom logic (e.g. perl, R)

- Integrating python meta-packages with full spack package
  management allows building optimized numpy stack AND easy python management.
\end{verbatim}