### 2015-04-18 usecases.tex

%!TEX root = paper.tex

\section{Use Cases}
\label{sec:usecases}

We have recently using Spack in production at 
Livermore Computing.  We have already outlined a number of the
experiences that drove us to develop Spack.
In this section, we describe real-world use cases that Spack
has been used to address.  In some cases, Spack had to be 
adapted to meet production needs, and we describe how its
flexibility has allowed us to put together solutions quickly.

\input{usecase-combinatoric}
\input{usecase-python}
\input{usecase-policy}