### 2015-04-18 usecases.tex
edits to usecases

%!TEX root = paper.tex

\section{Use Cases}
\label{sec:usecases}

We have recently begun using Spack in production at 
Livermore Computing.  We have already outlined a number of the
experiences that drove us to develop Spack.
In this section, we describe real-world use cases that Spack
has addressed.  In some cases, Spack had to be 
adapted to meet production needs, and we describe how its
flexibility has allowed us to put together solutions quickly.

\input{usecase-combinatoric}
\input{usecase-python}
\input{usecase-policy}