### 2015-04-18 usecase-combinatoric.tex
edits to usecase combo

%!TEX root = paper.tex

\subsection{Combinatoric Naming}
\label{sec:usecase-combinatoric}
%\subsubsection{Gperftools and MPILeaks}

{\tt Gperftools} is a suite of tools from Google that has gained popularity among
developers for its high-performance thread-safe heap and its lightweight profilers.
Unfortunately, two issues made it difficult to maintain {\tt gperftools} installations
at LLNL.  First, {\tt gperftools} is a C++ library.  Since C++ does not define a standard
ABI, {\tt gperftools} must be re-built with each compiler and compiler version used by client
applications.  Second, building {\tt gperftools} on bleeding-edge architectures
(such as Blue Gene/Q) requires patches and complicated configure lines that
change with each compiler.  One LLNL application team tried to maintain their own
builds of {\tt gperftools} for their preferred compilers, but this
task was complicated by the second issue.

Spack presented a solution to both problems.  Package administrators can use Spack to
easily maintain a central install of {\tt gperftools} across combinations of
versions and compilers.  Spack's {\tt gperftools} package also serves as a central
institutional knowledge repository for builds, as its package files encode
the patches and configure lines required for each platform and compiler combination.
We are investigating whether we can move some of these settings into compiler and
architecture descriptors to further simplify Spack's build template for cross-platform
installations.
\begin{figure}
\begin{minted}[linenos,
               numbersep=5pt,
			   fontsize=\scriptsize,
               frame=lines,
               framesep=2mm]{python}
patch("patch.gpeftools2.4_xlc", when="@2.4 %xlc")

def install(self, spec, prefix):
    if spec.architecture == "bgq" and self.compiler.satisfies("xlc"):
        configure("--prefix=" + prefix,  "LDFLAGS=-qnostaticlink")
    elif spec.architecture == "bgq":
        configure("--prefix=" + prefix,  "LDFLAGS=-dynamic")
    else:
        configure("--prefix=" + prefix)

    make()
    make("install")
\end{minted}
  \caption{
    Simplified install routine for gperftools.
    \label{fig:gperftools}
  }
\end{figure}

Figure~\ref{fig:gperftools} illustrates per-compiler and platform build rules with
a simplified version of the install routine for {\tt gperftools}.
%(the full {\tt gperftools} install routine includes other compilers and more options).
The build applies a patch if {\tt gperftools} 2.4 is built with the XL compiler,
and it selects the correct configure line based on the spec's platform and compiler.

LLNL's {\tt mpileaks} tool, which we have used as a running example in this paper,
has an even larger configuration space than {\tt gperftools}.  It must be built
for different compilers, compiler versions, MPI implementations, and MPI implementation
versions. As with {\tt gperftools}, we have been able to install many different
configurations of {\tt mpileaks} using Spack.  Moreover, Spack's virtual dependency system
allows us to
quickly compose a new {\tt mpileaks} build when a new MPI library is deployed at LLNL, {\it without}
modifying the {\tt mpileaks} package itself.
%Though we simplified its dependency structure
%here, the {\tt callpath} library can be built with three different stack trace libraries:
%{\tt dyninst}, {\tt libunwind}, or {\tt glibc}.  For some codes we may soon need to build
%different versions of {\tt mpileaks} with different {\tt callpath} configurations.