### 2015-04-18 implementation.tex

%!TEX root = paper.tex

\section{The Spack Package Manager}
\label{sec:implementation}
Based on our experiences at LLNL, we have developed
{\it Spack}, the Supercomputing PACKage manager.
Spack is written in Python.  We chose Python for its flexibility
and its wide use in the HPC community.
%
Like prior systems, Spack supports an arbitrary number of software
installations, and like Nix it can identify them with hashes.  Unlike any
prior system, Spack provides a language to specify and manage the
combinatorial space of HPC software configurations.

\noindent
Spack has the following unique capabilities:
\begin{enumerate}
\item Packages are {\bf parameterized}, allowing new builds to be composed
      on the fly for different versions, architecture, compiler, options, 
      and dependencies.
\item To manage this large parameter space, spack provides a novel, 
      recursive {\bf spec syntax} for dependency graphs.
\item Specs support versioned, ABI-incompatible interfaces like MPI through
      {\bf versioned virtual dependencies}.
\item Spack builds with {\bf compiler wrappers} that enforce build
      consistency and simplify package writing.
\end{enumerate}

\input{implementation-packages}
\input{implementation-specs}
\input{implementation-virtual-deps}
\input{implementation-concretization}
\input{implementation-environment}