### 2015-04-19 conclusion.tex
Actual submitted version

%!TEX root = paper.tex

\section{Conclusion}
\label{sec:conclusion}

The complexity of managing HPC software is rapidly increasing, and it
will continue unabated without better tools.
In this paper, we reviewed the state of software management tools 
across a number of HPC sites, with particular focus on Livermore
Computing (LC). While tools exist that can handle multi-configuration
installs, none of them addresses the combinatorial nature
of the software configuration space directly. None of them allows
a user to rapidly {\it compose} new parametric builds.

We introduced Spack, a package manager in development at LLNL, that
provides truly {\it parameterized} builds.  Spack implements
a novel, recursive {\it spec} syntax that simplifies the process of working
with large software configuration spaces, and it builds software 
so that it will run correctly, regardless of the environment.
We outlined a number of Spack's unique features, including
versioned virtual dependencies, and the {\it concretization} process,
which converts an underspecified build DAG into a buildable spec.

We showed through three use cases that Spack is already
increasing operational efficiency 
in production at LLNL.  We believe that the software management
techniques implemented in
Spack are applicable to a broad range of HPC facilities.
Spack is available online at: http://bit.ly/spack-git.