### 2015-06-12 rebuttal.txt
Added initial rebuttal draft.

We thank the reviewers for their thoughtful comments. We will fix all
grammatical errors and phrasing issues. Below we address the major
concerns of the reviewers.

Reviewer 2 asks for new scientific contributions. We first note that
this is a State of the Practice (SOP) submission. From the SC15
website:

    A SOP paper can describe a first-of-its-kind technology or
    methodology or can capture a unique perspective (or experience) on
    issues, challenge, and solutions for dealing with aspects of
    unprecedented scale and complexity, particularly the experiences
    and knowledge that can be generalized to wide ranges of systems
    and usages [...] how the experience generalizes to wider
    applicability should be explored.

    (http://sc15.supercomputing.org/program/technical-papers/papers-faq)

This paper matches this description, as well as these SOP topics from
the SC15 CFP:

  1. Deployment experiences of large-scale infrastructures &
     facilities.

  2. User support experiences with large-scale and novel machines.

Further, Spack offers novel research contributions. In particular:

  a. No other tool offers a concise, grammar to specify constraints on
     elements in a dependency DAG.

  b. A novel concretization process to translate the relaxed spec
     grammar into a full build specification.

  c. First-class support for versionsed dependencies, versioned
     virtual dependencies, versioned compilers, and build constraints.

To our knowledge, no other system gives the user this much control
over parameterized builds, and no other system handles the full
combinatoric complexity or constraint management for software
dependency DAGs on modern HPC systems.

Reviewer 2 asks for quantitative measures of success.  Using other
tools, it was impossible to automatically build an LLNL production
code in all supported configurations. Using Spack's parameterized
builds, we can build the code with 50-60 dependencies in 20
configurations on 2 architectures, nightly. Each such build replaces
what was previously a multi-day effort.

Reviewers 2 and 3 ask for limitations and downsides. Spack can require
more disk space than a purely environment-module-based system, as we
require otherwise identical packages with different dependencies to be
built separately. In our view, the significant reduction in complexity
for the end user makes this worthwhile.

Reviewer 2 asks for barriers to entry for end users. Recursive syntax
has not been a barrier. As designed, Spack requires only a package
name in most cases, and constraints allow specificity *when
needed*. End users with mid- or large-size projects have readily
adopted Spack.  Python has been a barrier for some and a major
attraction for others. For those who would otherwise create local
workarounds, Python's flexibility has allowed them to extend local
Spack packages to suit their needs. Such customizations frequently
guide our priorities for new core features.

Reviewer 3 asks for performance overhead. Spack commands run quickly;
concretization takes no more than seconds even for packages with many
dependencies. Spack builds run faster than typical manual builds
because spack automatically stages them in a fast tmp
filesystem. Overhead of our compiler wrapper is negligible.

Reviewer 3 mentions the ROCKS distribution, which formerly used RPM
and now uses Anaconda. Both are binary package managers and are cited
in Section 2. We will explicitly mention ROCKS in the final paper.

Reviewer 4 asks what remains for future work. Spack uses a greedy
constraint solver, which is sufficient for current packages. We may
need full SAT solving in the future. Architecture support is currently
LLNL-specific, and we plan to support a broader range of OS and system
types. Currently no system supports allows packages to depend on
particular compiler features, e.g. support for C++11 language features
or support for specific OpenMP versions. We plan to add this
capability to our already robust compiler support.

We will clarify all of the points above in the final paper.

Finally, Reviewer 1 mentions the need to grow a broader community for
Spack. We wholeheartedly agree. Outside LLNL, we have received
contributions from Argonne National Laboratory, Los Alamos National
Laboratory, Illinois Institute of Technology, Stowers Institute for
Medical Research, UC Davis, and U. Canterbury.  Via weekly
teleconferences and a public mailing list, we actively collaborating
with these organizations, as well as groups at Sandia and Intel.

We are actively working to grow the base of contributors, as a tool
like Spack will only survive in the long term with community support.
We submitted to SC15 for greater visibility in the HPC community, and
because SC is a well-known top-tier venue for HPC tools.