### 2015-07-19 implementation-specs.tex

%!TEX root = spack-paper-sc15.tex

\begin{figure}
	\begin{subfigure}{\linewidth}
		\centering
		\includegraphics[width=\columnwidth]{specs/mpileaks.pdf}
		\caption{
			Spec for {\tt mpileaks}
			\label{fig:specs-mpileaks}
		}
	\end{subfigure}
%
	\begin{subfigure}{\linewidth}
		\centering
		\includegraphics[width=\columnwidth]{specs/mpileaks-version}
		\caption{
			{\tt mpileaks@2.3}
			\label{fig:specs-mpileaks-version}
		}
	\end{subfigure}
%
	\begin{subfigure}{\linewidth}
		\centering
		\includegraphics[width=\columnwidth]{specs/mpileaks-abstract.pdf}
		\caption{
			{\tt mpileaks@2.3 \^{}callpath@1.0+debug \^{}libelf@0.8.11}
			\label{fig:specs-mpileaks-abstract}
		}
	\end{subfigure}
%
	\caption{
		Constraints applied to {\tt mpileaks} specs.
	}
	\label{fig:specs}
\end{figure}

\subsection{Spack Specs}\label{sec:specs}

Using the simple script in Figure~\ref{fig:mpileaks}, Spack can build many different
versions and configurations of the {\tt mpileaks} package.  In traditional port systems,
package code is structured to build a single version of a package, but in Spack, each
package file is a {\it template} that can be configured and built in many
different ways, according to a set of {\it parameters}.
Spack calls a single build configuration a {\it spec}. The {\tt spec} object
passed to {\tt install()} is how Spack communicates dependencies and parameters to package authors.

\subsubsection{Structure}
To understand how specs work, consider the {\tt mpileaks} package structure.
Metadata in the {\tt Mpileaks} class ({\tt version}, {\tt depends\_on}, etc.) describe
its relationships with other packages.  There are two direct dependencies:
the {\tt callpath} library and {\tt mpi}.  Spack recursively inspects the class definitions
for each dependency and constructs a graph of their relationships.  The result
is a directed, acyclic graph (DAG) of dependencies.\footnote{Spack currently disallows
circular dependencies.}
%
To guarantee a consistent build, and to avoid ABI compatibility issues, the DAG
is constructed so that there is only {\it one} version of any particular package.  Note
the distinction: while Spack can install arbitrarily many configurations of any package,
no two configurations of the same package will ever appear in the same build DAG.

DAGs for {\tt mpileaks} are shown in Figure~\ref{fig:specs}.
Each node represents a package, and each package has five configuration parameters that control
how it will be built: 1) the package version, 2) the compiler to
build with, 3) the compiler version, 4) named compile-time build options, or {\it variants},
and 5) the target architecture.

\subsubsection{Configuration Complexity}
A spec DAG has many degrees of freedom, and it is not reasonable to expect users to
understand or specify all of them.  In our experience at LLNL, the typical user
only cares about a small number of build constraints (if any), and does not know enough to
specify the rest. For example, a user may know that a certain version of a library like
{\tt boost}~\cite{boost} is required, but only cares that other build parameters are set so that
the build will succeed.
%
Configuration complexity makes the HPC software ecosystem difficult to manage: there are
simply too many parameters. However, the small set of important build constraints can be
very specific, so we have two competing concerns.  We need the ability to specify details
of the configuration, without the complexity of remembering all of it.

\subsubsection{Spec Syntax}\label{sec:syntax}
\begin{figure}
{\fontfamily{cmr}\selectfont
\begin{grammar}\small
  <spec>         ::= <id> [ <constraints> ]

  <constraints>   ::= \{ `@' <version-list> | `+' <variant> \newline
                   | `-' <variant> ~~~| `~' <variant> \newline
                   | `\%' <compiler> ~| `=' <architecture> \} \newline
                  [ <dep-list> ]

  <dep-list>  ::= \{ `\textsf \textasciicircum' <spec> \}

  <version-list> ::= <version> [ \{ `,' <version> \} ]

  <version>      ::= <id> | <id> `:' | `:' <id> | <id> `:' <id>

  <compiler>     ::= <id> [ <version-list> ]

  <variant>      ::= <id>

  <architecture> ::= <id>

  <id>           ::= "[A-Za-z0-9_][A-Za-z0-9_.-]*"
\end{grammar}
} % End fontfamily{cmr}
\caption{
	EBNF grammar for spec expressions.
	\label{fig:grammar}
}
\end{figure}

\begin{table*}\centering
\begin{tabular}{|r|p{2.4in}|p{4in}|}
\hline
& {\bf Spec} & {\bf Meaning} \\
\hline
\hline
1&\small\verb|mpileaks|                         & \small {\tt mpileaks} package, no constraints. \\\hline
2&\small\verb|mpileaks@1.1.2|                   & \small {\tt mpileaks} package, version 1.1.2. \\\hline
3&\small\verb|mpileaks@1.1.2 %gcc|              & \small {\tt mpileaks} package, version 1.1.2, built with {\tt gcc} at the default version. \\\hline
4&\small\verb|mpileaks@1.1.2 %intel@14.1 +debug| & \small {\tt mpileaks} package, version 1.1.2, built with Intel compiler version 14.1, \newline with the ``debug'' build option. \\\hline
5&\small\verb|mpileaks@1.1.2 =bgq|              & \small {\tt mpileaks} package, version 1.1.2, built for the Blue Gene/Q platform (BG/Q). \\\hline
6&\small\verb|mpileaks@1.1.2 ^mvapich2@1.9|     & \small {\tt mpileaks} package version 1.1.2, using {\tt mvapich2}  version 1.9 for MPI. \\\hline
7&\small\verb|mpileaks @1.2:1.4 %gcc@4.7.5 -debug =bgq| \newline
      \verb|  ^callpath @1.1 %gcc@4.7.2| \newline
      \verb|  ^openmpi @1.4.7|                & \small%
      {\tt mpileaks} at any version between 1.2 and 1.4 (inclusive), built with gcc 4.7.5,
      without the debug option, for BG/Q, linked with {\tt callpath} version 1.1
      and building {\tt callpath} with {\tt gcc} version 4.7.2, linked with {\tt openmpi} version 1.4.7.    \\
\hline
\end{tabular}
\caption{
	Spack build spec syntax examples and their meaning.
	\label{tab:specs}
}
\end{table*}

We have developed a syntax for specs that allows users to specify
only those constraints they care about, but to be specific when necessary.
Our syntax is expressive enough to represent software DAGs  but concise enough to use
on the command line. The spec syntax is recursively defined to allow users to specify
parameters on dependencies as well as on the root of the DAG.  The EBNF grammar we use
to implement this in Spack is shown in Figure~\ref{fig:grammar}.

We begin with a simple example.
Consider the case where a user wants to install the {\tt mpileaks} package, but knows
nothing about its structure.  To install the package, the user invokes {\tt spack install}:
%
\begin{minted}[fontsize=\scriptsize]{bash}
    $ spack install mpileaks
\end{minted}
%
Here, {\tt mpileaks} is the simplest possible spec---a single identifier.
Spack parses it and converts it into the DAG shown in Figure~\ref{fig:specs-mpileaks}.
Note that even though the spec contains no dependency information, it is still
converted to a full DAG, based on directives supplied in package files. Since there
are no constraints on the nodes, Spack has considerable leeway when building the
package, and we say that the package is {\it unconstrained}.
%
Now, suppose the user wants a specific version of {\tt mpileaks}.  This can be requested
with a version constraint after the package name:
%
\begin{minted}[fontsize=\scriptsize]{bash}
    $ spack install mpileaks@2.3
\end{minted}
%
We see from Figure~\ref{fig:specs-mpileaks-version} that the specific version constraint is
placed on the {\tt mpileaks} node in the DAG, but the rest of the DAG remains unconstrained.
If the user does not need a specific version of {\tt mpileaks}, but does require
particular minimum version, then the user could use {\it version range} syntax
and write {\tt @2.3:}.  Likewise, for a version between 2.3 and 2.5.6, she would use
{\tt @2.3:2.5.6} to designate the range. In these cases, the user can save time if
Spack already has a version installed that satisfies the spec -- Spack will use
the previously-built installation instead of building a new one.

Figure~\ref{fig:specs-mpileaks-abstract} shows the recursive nature of the spec syntax:
%
\begin{minted}[fontsize=\scriptsize]{bash}
    $ spack install mpileaks@2.3 ^callpath@1.0+debug ^libelf@0.8.11
\end{minted}
%
The caret (\verb|^|) denotes constraints for a particular dependency.  In the DAG,
we see that there are now version constraints on {\tt callpath} {\it and} {\tt libelf},
and that the user has requested the {\tt debug} variant of the {\tt callpath} library.

Recall that Spack guarantees that there is only a single version of any package in
the spec DAG.  Within the same DAG, each dependency can be uniquely
identified by its package name alone.  Therefore, the user need not think about DAG 
connectivity to add constraints.  She need only know that {\tt mpileaks}
somehow depends on {\tt callpath}.
For the same reason, the constraint order does not matter; dependency constraints
can appear in arbitrary order.

Table~\ref{tab:specs} shows further examples of specs, ranging from very
simple to more complex. From these examples, we can see that Spack offers constraint
notation to cover the rest of the HPC package parameter space.

{\bf Versions.}
Version constraints, denoted with {\tt @}, were already covered above. Versions
can be precise ({\tt @2.5.1}) or denote a range ({\tt @2.5:4.4}),
which may be open-ended ({\tt @2.5:}).

The package in Figure~\ref{fig:mpileaks} lists two ``safe'' versions with checksums, but
in our experience users frequently want bleeding-edge versions, and package managers
frequently lag behind the latest releases.
Spack has a capability to extrapolate URLs from versions,
using the package's {\tt url} attribute as a model.\footnote{This works
for packages with consistently named URLs.}  The user can request a specific
version on the command line, even if it is unknown to Spack,
and Spack will attempt to install it.  Spack also uses the same
model to scrape webpages and find new versions as they become available.

{\bf Compilers.}
With a compiler constraint (shown on line 3) the user
simply adds {\tt \%} followed by its name, along with an optional compiler version
specifier.  Spack compiler names, e.g. {\tt gcc}, refer to the full compiler {\it toolchain},
i.e. the C, C++, Fortran 77, and Fortran 90 compilers.  Spack can auto-detect
compiler toolchains if they are in the user's {\tt PATH}, or they can be registered manually
through a configuration file.

{\bf Variants.}
To handle build options like compiler flags or optional components, specs can
have named flags, or {\it variants}.  Variants are associated with the package,
so the {\tt mpileaks} package implementor must check the spec and handle the cases
where debug is enabled ({\tt +debug}) and disabled (with {\tt -debug}
or {\tt \textasciitilde \ignorespaces debug}).  The names simplify the versioning and prevent
Spack's configuration space from becoming too fine-grained.
It would violate our goal of conciseness, for example, to include detailed
compiler flags in spec syntax, but known sets of flags can simply be named.

{\bf Cross-compilation.}
To support cross-compilation, Spack includes the platform in the package spec (line 5).
Platforms begin with {\tt =} and take names like {\tt linux-ppc64} or {\tt bgq}.  They are
specified per-package; this allows front-end tools to depend on their back-end measurement
libraries with a {\it different} architecture on cross-compiled machines.

\subsubsection{Constraints in packages}\label{sec:constraints}

So far, we have shown examples of specs being used to request constraints from the
command line, when {\tt spack install} is invoked.  However, the user is not the only
source of constraints.  Applications may require specific versions of dependencies,
and it is often desirable to write constraints like this into a package file.  For
example, the ROSE compiler only builds with a certain version of the {\tt boost} library.
At first glance, the {\tt depends\_on()} directives in Figure~\ref{fig:mpileaks} look
like they take simple package names.  However, a package name is also a spec, and
the same constraint syntax usable from the command line can be applied inside directives.
So, the ROSE compiler can simply write:
%
\begin{minted}[fontsize=\scriptsize]{python}
    depends_on('boost@1.54.0')
\end{minted}
%
This constraint will be incorporated into the initial DAG node generated from
the {\tt ROSE} package.  Dependencies can also be {\it conditional}.  For example,
the {\tt ROSE} package builds with different versions of boost, depending on the
compiler version. To handle this, directives also accept spec syntax as a
predicate in an optional {\tt when} parameter:
%
\begin{minted}[fontsize=\scriptsize]{python}
    depends_on('boost@1.47.0', when='%gcc@:4')
    depends_on('boost@1.54.0', when='%gcc@5:')
\end{minted}
%
The same notation is used to optionally build with libraries like MPI:
%
\begin{minted}[fontsize=\scriptsize]{python}
    depends_on('mpi', when='+mpi')
\end{minted}
%
and to ensure that specific patches are applied to Python source code
when it is built on Blue Gene/Q, with different compilers:
%
\begin{minted}[fontsize=\scriptsize]{python}
    patch('python-bgq-xlc.patch',   when='=bgq%xl')
    patch('python-bgq-clang.patch', when='=bgq%clang')
\end{minted}
%
Constraints in the {\tt when} clause are matched against the package spec.
Outside of directives, constraints can also be tested directly on the {\tt spec}
object in the {\tt install} method:
%
\begin{minted}[fontsize=\scriptsize]{python}
  def install(self, spec, prefix):
      if spec.satisfies('%gcc'):
          # Handle gcc
      elif spec.satisfies('%xl'):
          # Handle XL compilers
      ...
\end{minted}
%

\subsubsection{Build specialization}

In our experience maintaining packages at LLNL, we often have to change
entire package build scripts due to large changes in the way certain packages build.
This can be very cumbersome, and it is difficult to maintain both the old and new
version of a build script, which we must do for users who rely on the old version.

%class Dyninst(Package):
%    """"Dyninst binary instrumentation and analysis tool."""
%    url="http://www.paradyn.org/release8.1/DyninstAPI-8.1.1.tgz"
%    version('8.1.1', 'd1a04e995b7aa70960cd1d1fac8bd6ac')
%    depends_on("libelf")
%    depends_on("libdwarf")
%    depends_on("boost@1.42:")
%
\begin{figure}
\begin{minted}[linenos,
               numbersep=5pt,
			   fontsize=\scriptsize,
               frame=lines,
               framesep=2mm]{python}
    def install(self, spec, prefix):  # default build uses cmake
        with working_dir('spack-build', create=True):
            cmake('..', *std_cmake_args)
            make()
            make("install")

    @when('@:8.1')                    # <= 8.1 uses autotools
    def install(self, spec, prefix):
        configure("--prefix=" + prefix)
        make()
        make("install")
\end{minted}
\caption{
    Specialized {\tt install} method in Dyninst.
	\label{fig:specialization}
}
\end{figure}

For cases like this, Spack provides functionality that allows Python functions to have
multiple definitions, each specialized for particular configurations of the package.
This allows us to have two separate implementations of {\tt install} or {\it any} method
in a package class. Figure~\ref{fig:specialization} shows how this is used in the Dyninst
package.  The {\tt @when} directive is a Python decorator: a higher order function that
takes a function definition as a parameter and returns a new function to replace it.
We replace the function with a callable multi-function dispatch object, and we
integrate the predicate check into the function dispatch mechanism.  Here, the
{\tt @when} condition is true when Dyninst is at version 8.1 or lower, and in those cases
the package will use the {\tt configure}-based build.  By default if no predicate matches,
{\tt install} will use the default CMake-based implementation. The simple {\tt @when}
annotation allows us to maintain our old build code alongside the new version without
accumulating complex logic in a single {\tt install} function.