### 2015-07-19 limitations.tex

%!TEX root = spack-paper-sc15.tex

\subsection{Limitations and Future Work}

In future work we will investigate adding a SAT solver to Spack's
concretization process to present users with better options.  In our
experience, however, complex policy interactions like these are very
rare.