### 2015-07-19 limitations.tex
Edits through use cases.

%!TEX root = spack-paper-sc15.tex

\subsection{Limitations and Future Work}

In future work we will investigate adding a SAT solver to Spack's
concretization process to present users with better options.  In our
experience, however, complex policy interactions like these are very
rare.

% compilers and architectures
We are investigating whether we can move some of these settings into compiler and
architecture descriptors to further simplify Spack's build template for cross-platform
installations.