### 2015-07-20 implementation.tex
clarify unique features.

%!TEX root = spack-sc15.tex

\section{The Spack Package Manager}
\label{sec:implementation}
Based on our experiences at LLNL, we have developed
{\it Spack}, the Supercomputing PACKage manager.
Spack is written in Python.  We chose Python for its flexibility
and its wide use in the HPC community.
%
Like prior systems, Spack supports an arbitrary number of software
installations, and like Nix it can identify them with hashes.  Unlike any
prior system, Spack provides a language to specify and manage the
combinatorial space of HPC software configurations.

\noindent
Spack provides the following unique features:
\begin{enumerate}
\item Packages explicitly {\bf parameterized} by version, platform,
      compiler, options, and dependencies, to allow easy composition.
\item A novel, recursive {\bf spec syntax} for dependency graphs and constraints,
      which aids in managing the build parameter space.
\item {\bf Versioned virtual dependencies} to handle versioned, 
      ABI-incompatible interfaces like MPI.
\item A novel {\bf concretization} process that translates an abstract build
      specification into a full, concrete build specification.
\item A build environment with {\bf compiler wrappers} that enforce build
      consistency and simplify package writing.
\end{enumerate}

\input{implementation-packages}
\input{implementation-specs}
\input{implementation-virtual-deps}
\input{implementation-concretization}
\input{implementation-environment}