### 2015-07-20 limitations.tex
Limitations section.

%!TEX root = spack-sc15.tex

\subsection{Limitations and Future Work}

There are currently several limitations to Spack as implemented.
As mentioned in Section~\ref{sec:concretization}, the concretization
algorithm makes choices greedily. For example, if a package {\tt P} depends on 
both {\tt hwloc@1.9} and {\tt mpi}, and if the concretization algorithm
chooses an MPI implementation for {\tt P} that depends strictly on
{\tt hwloc@1.8}, this will lead to a conflict on {\tt hwloc}, and Spack
will raise an error and the user must resolve the issue.
Our implementation will not backtrack to find an MPI version
that does not conflict. So far, cases like this have been rare.
However, we plan to add better constraint satisfaction solving to
Spack in future work.

While Spack supports different architectures as part of the 
configuration space, we do not currently have ways to factor common
preferences (like {\tt configure} arguments and architecture-specific
compiler flags) out of packages and into separate architecture
descriptions.  This leads to some clutter in the package files when
too many per-platform conditions accumulate.  We are in the process
of adding features that would further simplify Spack's build templates
for cross-platform installations.

Spack can require more disk space than a module-based
system, as we require otherwise identical packages with different
dependencies to be built separately. In our view, the significant
reduction in complexity for the end user justifies this cost.

The use of Python has been a barrier for some users, and
a major attraction for others. For those who would otherwise create local
workarounds, Python's flexibility has allowed them to extend local
Spack packages to suit their needs. Such customizations frequently
guide our priorities for new core features. We believe that the
benefits of using a popular, rapidly emerging HPC language for our
embedded DSL outweigh any costs of the learning curve.

In addition to the future directions already mentioned,
our main near-term goal is to grow a community around Spack and to
build a larger base of contributors.  On the technical side, we will
increase build robustness by deploying a continuous testing system
at LLNL.  Finally, to support the growing number of
HPC languages, runtimes, and programming models, we plan to add
capabilities to Spack that allow packages to depend on particular
compiler features.  More and more, our codes are beginning to rely
on advanced compiler capabilities, like {\tt C++11} language features,
OpenMP versions, and GPU compute capabilities. Ideally, Spack
will find suitable compilers and ensure ABI consistency when many
such features are in use.