### 2015-07-20 conclusion.tex
Build environment figure text.

%!TEX root = spack-sc15.tex

\section{Conclusion}
\label{sec:conclusion}

The complexity of managing HPC software is rapidly increasing, and it
will continue unabated without better tools.
In this paper, we reviewed the state of software management tools
across a number of HPC sites, with particular focus on Livermore
Computing (LC). While tools exist that can handle multi-configuration
installs, none of them addresses the combinatorial nature
of the software configuration space directly. None of them allows
a user to rapidly {\it compose} new builds with version, compiler,
and dependency parameters.

We introduced Spack, a package manager in development at LLNL, that
provides truly {\it parameterized} builds.  Spack implements
a novel, recursive {\it spec} syntax that simplifies the process of working
with large software configuration spaces, and it builds software
so that it will run correctly, regardless of the environment.
We outlined a number of Spack's unique features, including
versioned virtual dependencies, and its novel {\it concretization}
process, which converts an abstract build DAG into a concrete,
buildable spec.

We showed through four use cases that Spack is already increasing
operational efficiency in production at LLNL.  We believe that the
software management techniques implemented in Spack are applicable
to a broad range of HPC facilities.
Spack is available online at http://github.com/scalability-llnl/spack.