### 2015-07-20 introduction.tex
various small tweaks to intro

%!TEX root = spack-sc15.tex

\section{Introduction}
\label{sec:intro}

The Livermore Computing (LC) facility at Lawrence Livermore National Laboratory
(LLNL) supports around 2,500 users on 25 different clusters, ranging
in size from a 1.6 teraflop, 256-core cluster to the
20 petaflop, 1.6 million-core Sequoia machine, currently ranked first and
third on the Graph500~\cite{graph500} and Top500~\cite{top500}
lists, respectively.
%
%The massively parallel simulations that run on LC machines support
%U.S. Department of Energy missions across a range of scientific domains,
%including nuclear physics, material science, climate science, mechanical
%engineering, geophysics, seismology, and stockpile stewardship.
%
The simulation software that runs on these machines is very complex; some
codes depend on specific versions of over 50 dependency libraries.
They require specific compilers, build options and MPI implementations to
achieve the best performance, and users may run several
different codes in the same environment as part of larger
scientific workflows.

To support the diverse needs of applications, system administrators
and application developers frequently build, install, and support many
different configurations of math and physics libraries, as well as
other software.  Frequently, applications must be rebuilt to fix bugs
and to support new versions of the operating system (OS), Message
Passing Interface (MPI) implementation, compiler, and other
dependencies.  Unfortunately, building scientific software is
notoriously complex, with immature build systems that are difficult to
adapt to new
machines~\cite{dubois+:comp-sci-eng,hoste+:pyhpc12,wilson+:corr}.
Worse, the space of required builds grows combinatorially with each
new configuration parameter. As a result, LLNL staff spend countless
hours dealing with build and deployment issues.

Existing package management tools automate parts of the build
process~\cite{bsdports,digirolamo:smithy,dolstra+:icfp08,dolstra+:lisa04,hashdist,homebrew,hoste+:pyhpc12,macports,thiruvathukal:gentoo04}.
For the most part, these systems focus on keeping a single, stable set of
packages up to date, and they do not handle installation of multiple
versions or configurations.  Those that {\it do} handle multiple configurations
typically require that package files be created for each combination of
options~ \cite{digirolamo:smithy,dolstra+:icfp08,dolstra+:lisa04,hashdist,hoste+:pyhpc12},
leading to a profusion of files and maintenance issues.
To our knowledge, no existing tool allows new configurations to be composed
and assembled on demand.  Some allow limited forms of
composition~\cite{hoste+:pyhpc12,dolstra+:icfp08,dolstra+:lisa04}, but their
dependency management is overly rigid, and they burden users with
combinatorial naming and versioning problems.

This paper describes our experiences with the {\it Spack} package manager,
which we have developed at LLNL to manage increasing software complexity.
It makes the following contributions:
\begin{enumerate}
\item A novel, recursive syntax for concisely specifying and querying
      the large parameter space of HPC packages;
\item A build methodology that ensures packages find their dependencies
      regardless of users' environments;
\item Spack: An implementation of these concepts; and
\item Four use cases that detail LLNL's use of Spack to comcpose
      software configurations rapidly, to manage Python installations, and
      to implement site build policies, and to build an LLNL production code.
\end{enumerate}

\noindent Spack is in active development at LLNL, and our use cases outline some of
its first applications in our environment.  Spack solves software problems
that are pervasive at large, multi-user HPC centers, and our experiences
are relevant to the full range of HPC facilities.  Spack improves
operational efficiency by simplifying the build and deployment of
bleeding-edge scientific software.