### 2015-07-28 motivation.tex
Address shepherd comments through motivatoin.

%!TEX root = spack-sc15.tex

\section{Common Practice}
\label{sec:motivation}

\paragraph{Meta-build Systems}
{\it Meta-build systems} such as Contractor, WAF, and
MixDown~\cite{amundson:contractor,epperly+:mixdown,epperly+:mixdown-report,nagy:waf} are
related to package managers, but they focus on ensuring that a single
package builds with its dependencies.  MixDown notably provides excellent
features for ensuring consistent compiler flags in a build.
However, these systems do not provide facilities to manage
large package repositories or combinatorial versioning.

\paragraph{Traditional package managers}
Package managers automate the installation of complex sets of software packages.
{\it Binary package managers} such as RPM, yum, and APT~\cite{foster+:rpm03,silva:apt01,yum} are integrated with most
OS distributions, and they are used to ensure that dependencies
are installed before packages that require them.
Anaconda~\cite{anaconda,conda}, uses the same approach but
is designed to run on top of a host OS.
These tools largely solve the problem of managing a {\it single} software
stack, which works well for the baseline OS and drivers, which are
common to all applications on a system.
These tools assume that each package has only a single version,
and most install packages in a single, inflexible location.
To install multiple configurations, users must create custom, combinatorial
naming schemes to avoid conflicts. They typically require root
privileges and do not optimize for specific hardware.

{\it Port systems} such as Gentoo, BSD Ports, MacPorts, and
Homebrew~\cite{bsdports,groffen:gentoo-prefix,homebrew,macports,thiruvathukal:gentoo04}
build packages from source instead of installing from a pre-built binary.
Most port systems suffer from
the same versioning and naming issues as traditional package managers.
Some allow multiple versions to be installed in the same
prefix~\cite{groffen:gentoo-prefix}, but again the burden is on package
creators to manage conflicts. This burden effectively restricts installations
to a few configurations.

\paragraph{Virtual Machines and Containers}

Packaging problems arise in HPC because a supercomputer's hardware, OS, and
filesystem are shared by many users with different requirements.  The classic
solution to this problem is to use virtual
machines (VMs)~\cite{barham2003xen,rosenblum1999vmware,smith2005architecture}
or lightweight virtualization techniques like Linux
containers~\cite{felter2014updated,merkel2014docker}. This model allows each
 user to have a personalized environment with its own package manager, and it
has been extremely successful for servers at cloud data centers. VMs typically
have near-native compute performance but low-level HPC network drivers still
exhibit major performance issues. VMs are not well supported on many
non-Linux operating systems, an issue for the lightweight
kernels of bleeding-edge Blue Gene/Q and Cray machines.
Finally, each VM still uses a traditional package manager,
so running many configurations still requires a large number of VMs.
For facilities, this profusion of VMs is a security concern since it 
complicates mandatory patching. Also, managing  a large number of VM 
environments is tedious for users.

\begin{table*}\centering
\begin{tabular}{|l|l|}
\hline
Site           & Naming Convention \\
\hline
\hline
LLNL       & {\tt / usr / global / tools / \$arch / \$package / \$version} \\
           & {\tt / usr / local~ / tools / \$package-\$compiler-\$build-\$version } \\
\hline
Oak Ridge~\cite{jones+:cug08}  & {\tt / \$arch / \$package / \$version / \$build} \\
\hline
TACC/Lmod~\cite{mclay:lmod-tutorial}& {\tt / \$compiler-\$comp\_version / \$mpi / \$mpi\_version / \$package / \$version} \\
\hline
\hline
Spack default                  & {\tt / \$arch / \$compiler-\$comp\_version / \$package-\$version-\$options-\$hash} \\
\hline
\end{tabular}
\caption{
	Software organization of various HPC sites.
	\label{tab:naming-conventions}
}
\end{table*}

\paragraph{Manual and Semi-automated Installation}

To cope with software diversity, many HPC sites use a combination of existing
package managers and either manual or semi-automated installation.
For the baseline OS, many sites maintain traditional binary
packages using the vendor's package manager. LLNL maintains a Linux
distribution, CHAOS~\cite{chaos} for this purpose, which is managed using RPM.
The popular ROCKS~\cite{rocks} cluster distribution uses RPM and Anaconda
in a similar fashion.
%
For custom builds, many sites adhere to detailed naming conventions
that encode information in file system paths.
Table~\ref{tab:naming-conventions} shows several sites' conventions.
LLNL uses the APT package manager for installs
in the {\tt /usr/local/tools} file system and {\tt /usr/global/tools}
for manual installs.
Oak Ridge National Laboratory (ORNL) uses hand installs but adheres
to strict scripting conventions
to reproduce each build~\cite{jones+:cug08}.
The Texas Advanced Computing Center (TACC) relies heavily on locally maintained RPMs.

From the conventions in Table~\ref{tab:naming-conventions},
we see that most sites use some combination of architecture, compiler version,
package name, package version, and a custom (up to the author, sometimes
encoded) build identifier.  TACC and many other sites also explicitly
include the MPI version in the path. MPI is explicitly called out
because it is one of the most common software packages for HPC.
However, it is only one of many dependencies that go into a build.
None of these naming conventions covers the entire configuration
space, and none has a way to represent, e.g., two builds that are identical
save for the version of a particular dependency library.  In our experience
at LLNL, naming conventions like these have not succeeded because
users want more configurations than we can represent with a practical
directory hierarchy. Staff frequently install nonconforming packages
in nonstandard locations with ambiguous names.

\paragraph{Environment Modules and RPATHs}\label{sec:env-rpath}

Diverse software versions not only present problems for build and installation;
they also complicate the runtime environment. When launched, an executable
must determine the location of its dependency libraries, or it will not run.
Even worse, it may find the wrong dependencies and subtly produce incorrect results.
Statically linked binaries do not have this issue, but modern
operating systems make extensive use of dynamic linking.
By default, the dynamic loader on most systems is configured to search only
system library paths such as {\tt /lib}, {\tt /usr/lib}, and
{\tt /usr/local/lib}.  If binaries are installed in other locations, the
{\it user} who runs the program must typically add dependency library paths to
{\tt LD\_LIBRARY\_PATH} (or a similar environment variable) so that the loader
can find them.  Often, the user is not the same person who installed the library,
and even advanced users may have difficulty determining which paths to add.

Many HPC sites address this problem using {\it environment modules}, which
allow users to ``load'' and ``unload'' such settings dynamically using simple
commands. Environment modules emerged in 1991, and there are many implementations~\cite{dotkit,furlani+:lisa91,furlani+:lisa96,mclay:lmod,mclay:lmod-tutorial}.
The most advanced of these, Lmod~\cite{mclay:lmod,mclay:lmod-tutorial},
provides software hierarchies similar to the naming conventions in
Table~\ref{tab:naming-conventions}. This allows users to load a software stack
quickly if they know which one is required.

The alternative to per-user environment settings is to embed library search
paths in installed binaries at compile time. When set this way, the search
path is called an {\tt RPATH}. {\tt RPATHs} and environment modules are not
mutually exclusive. Modules can still be used to set variables that are
unrelated to linking, such as {\tt MANPATH} and {\tt PATH}.  Adding
{\tt RPATHs} still ensures that binaries run correctly,
regardless of whether the right module is loaded. LC installs software with
both {\tt RPATHs} and {\tt dotkit}~\cite{dotkit} modules.

\paragraph{Modern Package Managers}

Recently, a number of HPC package managers have emerged that manage
multi-configuration builds.
%
ORNL uses the Smithy~\cite{digirolamo:smithy} installation tool. It
can generate module files, but it does not provide any
automated dependency management; it only checks whether a package's
prerequisites have already been installed by the user.

The Nix~\cite{dolstra+:icfp08,dolstra+:lisa04} package manager and OS 
distribution supports installation of arbitrarily many software 
configurations.  As at most HPC sites, it installs each package
in a unique prefix but it does not have a human-readable naming 
convention.  Instead, Nix determines the prefix by hashing the package
file and its dependencies.
%% BRONIS: I'm not sure what the following sentence really adds
%%  Nix package files are written in a custom
%% functional language designed for packaging.

The EasyBuild~\cite{hoste+:pyhpc12} tool is in production use at
the University of Ghent and the J\"ulich Supercomputing Center.
It allows multiple versions to be installed at once.  Rather
than setting {\tt RPATHs}, it auto-generates module files
to manage the environment, and it is closely coupled with
Lmod~\cite{geimer+:hust14}.  EasyBuild groups the compiler, MPI, FFT, and
BLAS libraries together in a {\it toolchain} that can be used by
package files. This provides some degree of composability and
separates compiler flags and MPI concerns from client packages.

HashDist~\cite{hashdist} is a meta-build system and package manager
for HPC.  Of the existing solutions, it is the most similar to Spack.
Like Nix, it uses cryptographic versioning and stores installations in
unique directories.
%
Both Nix and HashDist use {\tt RPATHs} in their packages to ensure that
libraries are found correctly.

\paragraph{Gaps in current practice}
The flexible cryptographic versioning of Nix and HashDist manages 
the package {\it and} its dependency configuration and can represent any
configuration. However, users cannot easily navigate or query the installed
software.
%The systems are in a sense ``write-only'':
%Nix does not offer a way to view a package's dependencies.
%
EasyBuild and Smithy generate environment modules, which supports some 
querying. Naming schemes used in existing module systems, however, cannot 
handle combinatorial versions, which the Lmod authors call the 
``matrix problem''~\cite{mclay:lmod-tutorial}.

%EasyBuild's attempts to version groups dependencies by adding versions to
%toolchains, but the naming is difficult to understand.
%. Compiler, MPI, and
%some library versions are lumped together, but the results is cryptic, e.g.
%{\tt goolf} stands for ``gcc, openmpi, openblas, ScaLAPACK, FFTW'', and
%its version is meaningless.

%Existing tools do not enforce {\tt RPATHs}, leaving them to package authors.
%This may lead to erroneous runs.

The main limitation of existing tools is the lack of build {\it composability}.
The full set of package versions is combinatorial, and arbitrary combinations 
of compiler, MPI version, build options, and dependency versions require 
non-trivial modifications to many package files. Indeed, the number of package 
files required for most existing systems scales with the number of version 
{\it combinations}, not the number of packages, which quickly becomes 
unmanageable.  As an example, the EasyBuild system has over 3,300 files for 
several permutations of only 600 packages. A slightly different dependency 
graph requires an entire new package file hierarchy.  HashDist supports 
composition more robustly but does not have first-class parameters for 
versions, compilers, or versioned interfaces.
%
HPC sites need better ways to {\it parameterize} packages so that new
builds can be {\it composed} in response to user needs.
%