### 2015-07-31 limitations.tex

%!TEX root = spack-sc15.tex

\subsection{Limitations and Future Work}

Spack's current implementation has several limitations.
As mentioned in Section~\ref{sec:concretization}, we use a greedy 
concretization algorithm. For example, if package {\tt P} depends on 
both {\tt hwloc@1.9} and {\tt mpi}, and if the algorithm
chooses an MPI implementation for {\tt P} that depends strictly on
{\tt hwloc@1.8}, a conflict on {\tt hwloc} arises. In this case, Spack 
raises an error and the user must resolve the issue. Our implementation 
does not backtrack to find an MPI version that does not conflict. These 
cases have been rare so far. However, we plan to add better constraint
satisfaction solving to Spack in future work.

While Spack supports different architectures as part of the 
configuration space, we cannot currently factor common preferences (like 
{\tt configure} arguments and architecture-specific compiler flags) out of 
packages and into separate architecture descriptions, which leads to some 
clutter in the package files when too many per-platform conditions accumulate.
We are adding features that will further simplify Spack's build templates
for cross-platform installations.

Spack requires more disk space than a module-based system, as otherwise 
identical packages with different dependencies must be built separately. 
The exact space overhead depends on the structure of the installed
software; some builds can share more dependency libraries than others
(see Figure~\ref{fig:reuse}).
In our view, the significant reduction in complexity for the end user
justifies this cost.

The use of Python has been a barrier for some users, and
a major attraction for others. For those who would otherwise create local
workarounds, Python's flexibility has allowed them to extend local
Spack packages to suit their needs. Such customizations frequently
guide our priorities for new core features. Some users have found the 
learning curve to be too steep, but this only prevents them from 
packaging software. They still use Spack as a command-line tool.

In the near term, we are actively working to grow a community around Spack 
and to build a larger base of contributors. We plan to
increase build robustness by deploying a continuous testing system
at LLNL.  Finally, to support the growing number of
HPC languages, runtimes, and programming models, we will add
capabilities to Spack that allow packages to depend on particular
compiler features.  More and more, our codes are relying
on advanced compiler capabilities, like {\tt C++11} language features,
OpenMP versions, and GPU compute capabilities. Ideally, Spack
will find suitable compilers and ensure ABI consistency when many
such features are in use.