### 2015-08-10 introduction.tex

%!TEX root = spack-sc15.tex

\section{Introduction}
\label{sec:intro}

The Livermore Computing (LC) facility at Lawrence Livermore National Laboratory
(LLNL) supports around 2,500 users on 25 different clusters, ranging
in size from a 1.6 teraflop, 256-core cluster to the
20 petaflop, 1.6 million-core Sequoia machine, currently ranked second on the 
Graph500~\cite{graph500} and third on the Top500~\cite{top500}.
%
%The massively parallel simulations that run on LC machines support
%U.S. Department of Energy missions across a range of scientific domains,
%including nuclear physics, material science, climate science, mechanical
%engineering, geophysics, seismology, and stockpile stewardship.
%
The simulation software that runs on these machines is very complex; some
codes depend on specific versions of over 70 dependency libraries.
They require specific compilers, build options and Message Passing Interface (MPI)
implementations to
achieve the best performance, and users may run several
different codes in the same environment as part of larger
scientific workflows.

To support the diverse needs of applications, system administrators
and developers frequently build, install, and support many
different configurations of math and physics libraries, as well as
other software.  Frequently, applications must be rebuilt to fix bugs
and to support new versions of the operating system (OS), 
MPI implementation, compiler, and other
dependencies.  Unfortunately, building scientific software is
notoriously complex, with immature build systems that are difficult to
adapt to new
machines~\cite{dubois+:comp-sci-eng,hoste+:pyhpc12,wilson+:corr}.
Worse, the space of required builds grows combinatorially with each
new configuration parameter. As a result, LLNL staff spend countless
hours dealing with build and deployment issues.

Existing package management tools automate parts of the build
process~\cite{bsdports,digirolamo:smithy,dolstra+:icfp08,dolstra+:lisa04,hashdist,homebrew,hoste+:pyhpc12,macports,thiruvathukal:gentoo04}.
For the most part, they focus on keeping a single, stable set of
packages up to date, and they do not handle installation of multiple versions or
configurations.  Those that {\it do} handle multiple configurations
typically require that package files be created for each combination of
options~ \cite{digirolamo:smithy,dolstra+:icfp08,dolstra+:lisa04,hoste+:pyhpc12},
leading to a profusion of files and maintenance issues.
Some allow limited forms of 
composition~\cite{hoste+:pyhpc12,dolstra+:icfp08,dolstra+:lisa04}, but their
dependency management is overly rigid, and they burden users with
combinatorial naming and versioning problems.

This paper describes our experiences with the {\it Spack} package manager,
which we have developed at LLNL to manage increasing software complexity.
It makes the following contributions:
\begin{enumerate}
\item A novel, recursive syntax for concisely specifying constraints within
      the large parameter space of HPC packages;
\item A build methodology that ensures packages find their dependencies
      regardless of users' environments;
\item Spack: An implementation of these concepts; and
\item Four use cases detailing LLNL's use of Spack in production.
\end{enumerate}
\noindent
Our use cases highlight Spack's ability to manage complex software.
%
Spack supports rapid composition of package configurations, management 
of Python installations, and site-specific build policies. 
%
It automates 36 different build configurations of an LLNL
production code with 46 dependencies.
Despite this complexity, Spack's {\it concretization} algorithm
for managing constraints runs in seconds, even for
large packages. Spack's install environment incurs only around
10\% build-time overhead compared to a native install.

Spack solves software problems that are pervasive at large,
multi-user HPC centers, and our experiences are relevant to the
full range of HPC facilities.
Spack improves operational efficiency by simplifying the
build and deployment of bleeding-edge scientific software.