### 2015-08-10 usecase-python.tex

%!TEX root = spack-sc15.tex

\subsection{Support for Interpreted Languages}
\label{sec:usecase-python}

Python is becoming increasingly popular for HPC applications,
due to its flexibility as a language and its excellent support
for calling into fast, compiled numerical libraries.
Python is an interpreted language, but one can use it
as a friendlier interface to compiled libraries like FFTW, ATLAS, and
other linear algebra libraries.  Many LLNL code teams use Python in this manner.
%
LC supports Python installations for several application teams.
Maintaining these repositories has grown increasingly complex over
time. The problems are similar to those that drove us to create
Spack: different teams want different Python libraries with different
configurations.

Existing Python package managers either do not support building from
source~\cite{anaconda,conda}, or they are language-specific~\cite{eby:setuptools}.
None handles multi-configuration builds.  More glaringly, Python extensions
are usually installed into the Python interpreter's prefix, which makes it 
impossible to install multiple versions.\footnote{{\tt setuptools}
has support for multiple versions via {\tt pkg_resources},
but this requires modifications to client code.}
Installing each extension in its own prefix enables combinatorial versioning,
but it requires users to add packages to the {\tt PYTHONPATH} variable at runtime.

Per Spack's design philosophy, we wanted a way to manage many different
versions easily, but {\it also} to provide a baseline set of extensions 
{\it without} requiring environment settings.
%
To support this mode of operation, we added the concept of {\tt extension} packages
to Spack. Python modules use the \verb|extends('python')| directive instead of
\verb|depends_on('python')|.
Each module installs into its own prefix like any other package,
and each module depends on a particular Python installation.
However extensions can be {\tt activated} or {\tt deactivated}
within the dependent Python installation.  The {\tt activate} operation
symbolically links each file in the extension prefix into the Python
installation prefix, as if it were installed directly. If any file
conflict would arise from this operation, {\tt activate} fails.
Similarly, the {\tt deactivate} operation removes the symbolic links and restores
the Python installation to its  pristine state.

Additional complications arose because many Python packages 
{\it install their own package manager} if they do not find one in the 
Python installation. Also, Python packages use a wide range of mechanisms
to add themselves to the interpreter's default path, some of which conflict. 
We modified Spack so that extendable packages, like Python, can supply 
custom code in the package file that specializes {\tt activate} and 
{\tt deactivate} for the particular package. For Python, this feature 
merges conflicting files during activation. Overall, Python extensions can
be installed automatically in their own prefixes, and they can be composed
with a wide range of bleeding-edge libraries that other package managers do
not handle.

%To experiment with these extensions, users can load environment
%modules generated by Spack. If they want a particular version to be available
%without any special environment settings, they can activate it within the Python instance.

Spack essentially implements a ``meta package-manager'' for each Python
instance, which coexists with Spack's normal installation model.
This mechanism efficiently supports application teams, for whom we can
now rapidly construct custom Python installations.  We have also reduced
the time that LC staff spend installing Python modules.
Because Spack packages can extend activation and deactivation mechanisms,
this design could also be used with other languages
with similar extension models, such as R, Ruby, or Lua.